Stack Safety is a Security Property

### 2021-04-08 tex/main.tex
A comment and a few nits

The call stack is a perennial target for attacks
based on memory-safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}
\bcp{Sure, but later!}
\fi
This last provides an array of micro-policies offering protection at various granularities,
which offer a trade-off between protection and efficiently.

But what {\em is} the stack safety that these techniques enforce? Precise formal
definitions are mostly lacking: rather, the notion is commonly defined in a
negative way, by enumerating the bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
{\em local state encapsulation} and {\em well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft an operational characterization
of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
In other words, we seek formal criteria that can be applied
to a mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.'' We begin with a strong
local state encapsulation property that closely models the behavior of eager
micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
is too strong to characterize stack safety; reasonable enforcement mechanisms
exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
to support this technique and argue that the new property, defined relative to an
arbitrary notion of observation. This property is still strong enough to prevent
the bad behaviors we associate with unsafe stack behavior, but weak enough that
other reasonable enforcement mechanisms can enforce it.

@@

To phrase LSE and WBCF as security properties requires an explicit attacker model:
a clear understanding of who is being protected and from what. Most existing
stack protection mechanisms envisage scenarios where a callee is
influenced by external inputs to corrupt the stack in some way. To
avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we
state the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns. Thus we have no need to differentiate safe code
from adversarial - all functions, and indeed all activations of the same function,
are mutually distrusting.
%
For this model to make sense, we need to know when a call or a return is taking
place. We assume that calls and returns are annotated in the code, so that
we can track the point at which control transfers from the protected caller
to the attacking callee -- and when that callee in turn becomes a protected
caller in its own right.

@@

First we introduce this model for a single simple stack with no sharing between callers and
callee. Then we expand it to allow passing of stack data, granting a callee one time
access to the caller's data. We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. (A tag-based approach can easily
compose multiple such policies.)

>>>>

The call stack is a perennial target for attacks
based on memory-safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}
\bcp{Sure, but later!}
\fi
This last provides an array of micro-policies offering protection at various granularities,
which offer a trade-off between protection and efficiency.

But what {\em is} the stack safety that these techniques enforce? Precise formal
definitions are mostly lacking: rather, the notion is commonly defined in a
negative way, by enumerating the bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
{\em local state encapsulation} and {\em well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft an operational characterization
of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
In other words, we seek formal criteria that can be applied
to a mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.''
\rb{The remainder of this paragraph may be diving too quickly, would a reader be
confused by the references to these types of properties, yet to be introduced?}
We begin with a strong
local state encapsulation property that closely models the behavior of eager
micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
is too strong to characterize stack safety; reasonable enforcement mechanisms
exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
to support this technique and argue that the new property, defined relative to an
arbitrary notion of observation. This property is still strong enough to prevent
the bad behaviors we associate with unsafe stack behavior, but weak enough that
other reasonable enforcement mechanisms can enforce it.

@@

To phrase LSE and WBCF as security properties requires an explicit attacker model:
a clear understanding of who is being protected and from what. Most existing
stack protection mechanisms envisage scenarios where a callee is
influenced by external inputs to corrupt the stack in some way. To
avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we
state the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns. Thus we have no need to differentiate safe code
from adversarial --- all functions, and indeed all activations of the same function,
are mutually distrusting.
%
For this model to make sense, we need to know when a call or a return is taking
place. We assume that calls and returns are annotated in the code, so that
we can track the point at which control transfers from the protected caller
to the attacking callee --- and when that callee in turn becomes a protected
caller in its own right.

@@

First we introduce this model for a single simple stack with no sharing between callers and
callees. Then we expand it to allow passing of stack data, granting a callee one-time
access to the caller's data. We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. (A tag-based approach can easily
compose multiple such policies.)
